## At-Home Exercises

```{r h4_setup, echo = FALSE, include = FALSE}
source(here::here("code", "supportFunctions.R"))
```

This week, we will take our first steps into the world of factor analysis. We'll begin by fitting some CFA models using
personality scale data.

------------------------------------------------------------------------------------------------------------------------

### Setup

**Data:**

You will use the following dataset for these exercises.

- [cattell](https://surfdrive.surf.nl/files/index.php/s/1hLlshF9xF0rpJH)
- You can find a codebook for the dataset [here](data/16PF/codebook.html)

This dataset contains 1000 observations of 16X personality items collected as part of the
[International Personality Item Pool][ipip]. These items are meant to assess Raymond Cattell's 16 personality factors
(16PF):

- Warmth
- Reasoning
- Emotional Stability
- Dominance
- Liveliness
- Rule-Consciousness
- Social Boldness
- Sensitivity
- Vigilance
- Abstractness
- Privateness
- Apprehension
- Openness to Change
- Self-Reliance
- Perfectionism
- Tension

The [16PF Wikipedia page][cattell_wiki] provides a nice summary of the scale. 

**Packages:**

To complete these exercises, you will need to install the package described below.

```{r, echo = FALSE}
data.frame(
  Package = c(
    "semPlot"
  ),
  Description = c(
    "Programmatically draw path diagrams in R"
  )
) |>
  knitr::kable()
```

Use the `install.packages()` function to install this package.

<details>
  <summary>Click to show code</summary>

```{r, eval = FALSE}
install.packages("semPlot", dependencies = TRUE)
```

</details>

------------------------------------------------------------------------------------------------------------------------

###

Load the *cattell* data.

- The relevant data are contained in the [*canttell.rds*][cattell_data] file.

<details>
  <summary>Click to show code</summary>

```{r}
dataDir <- "data"
cattell <- readRDS(here::here(dataDir, "cattell.rds"))
```

</details>

------------------------------------------------------------------------------------------------------------------------

We'll first estimate a simple one-dimensional CFA of the *Warmth* factor. Our target model is defined as follows:

- One reflective latent factor representing the *Warmth* dimension
- 10 observed indicators: {`A1`, `A2`, ..., `A10`}
- 10 residual variances but no residual covariances

Basically, our CFA should estimate the default measurement model through which one latent factor generates 10 observed
indicator variables.

------------------------------------------------------------------------------------------------------------------------

###

Sketch a path diagram of the CFA model described above.

- Use pencil and paper (or the equivalent); don't generate the diagram programmatically.

------------------------------------------------------------------------------------------------------------------------

### {#df}

Calculate the following quantities for the *Warmth* CFA described above.

1. The number of parameters in the model.
1. The pieces of unique information provided by the data.
1. The number of estimated parameters after applying the necessary scaling constraints.
1. The degrees of freedom for the estimated model.

<details>
  <summary>Click to show the answer</summary>

1. The model contains 21 parameters
   - 10 factor loadings
   - 10 residual variances
   - One latent variance
1. Our model will include 10 observed variables, so the data provide 10 (10 + 1) / 2 = 55 pieces of information.
1. To identify the covariance structure model, we need to fix one parameter. So, we will estimate 20 parameters.
1. The estimated model will have 55 - 20 = 35 degrees of freedom.

</details>

------------------------------------------------------------------------------------------------------------------------

### {#cfaSyntax1}

Define the `lavaan` model syntax for the unidimensional CFA of *Warmth* described above.

- Do not specify any mean structure.
- Save this model syntax as an object in your environment.

<details>
  <summary>Click to show code</summary>

```{r}
mod1 <- '
warmth =~ A1 + A2 + A3 + A4 + A5 + A6 + A7 + A8 + A9 + A10
'
```

<details>
  <summary>Click for explanation</summary>

At this point, we only need to define the factor loading map. We can specify all the other options when we estimate the
model.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

### {#cfaEstimationHome1}

Estimate the CFA model you defined in \@ref(cfaSyntax1), and summarize the results.

- Use the `lavaan::cfa()` function to estimate the model.
- Use the default settings for the `cfa()` function.

Check the results, and answer the following questions:

- What is the estimated variance of the *Warmth* factor?
- How is the model identified when using the default settings?

<details>
  <summary>Click to show code</summary>

```{r}
## Load the lavaan package:
library(lavaan)

## Estimate the CFA model:
outH1.1 <- cfa(mod1, data = cattell)

## Summarize the fitted model:
summary(outH1.1)
```

<details>
  <summary>Click for explanation</summary>

The estimated latent variance is $\psi_{11} = `r inspect(outH1.1, "est")[["psi"]] |> as.numeric() |> round(3)`$.

The `cfa()` function is just a wrapper for the `lavaan()` function with several options set at the defaults you would
want for a standard CFA. 

- By default, the model is identified by fixing the first factor loading of each factor to 1 (due to the argument
`auto.fix.first = TRUE`).

To see a full list of the (many) options you can specify to tweak the behavior of `lavaan` estimation functions run 
`?lavOptions`.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

###

Do the number of model parameters and degrees of freedom reported in the lavaan output agree with the values you
calculated in \@ref(df)?

<details>
  <summary>Click to show answer</summary>

Yes, the values should agree. This first part of the lavaan model summary reports the number of model parameters and the
degrees of freedom.

```{r, echo = FALSE}
partSummary(outH1.1, 1:4)

fit <- fitMeasures(outH1.1)
p   <- fit[["npar"]] |> as.numeric() |> round()
df  <- fit[["df"]] |> as.numeric() |> round()
```

In this case, we see that we've estimated `r p` parameters and we have df = `r df`.

</details>

------------------------------------------------------------------------------------------------------------------------

###

Use the `semPaths()` function from the **semPlot** package to draw a path diagram of your estimated CFA.

- Include the parameter estimates in the diagram.

<details>
  <summary>Click to show code</summary>

```{r}
library(semPlot)
semPaths(outH1.1, whatLabels = "estimates")
```
</details>

------------------------------------------------------------------------------------------------------------------------

Now, we're going to complicate matters slightly by adding another dimension to our model. We'll now estimate a
two-dimensional CFA that contains correlated *Warmth* and *Dominance* factors.

Defined this new model as follows:

- Two correlated, reflective latent factors: one *Warmth* factor and one *Dominance* factor
- 10 observed indicators of *Warmth*: {`A1`, `A2`, ..., `A10`}
- 10 observed indicators of *Dominance*: {`D1`, `D2`, ..., `D10`}
- No cross-loadings
- No residual correlations

Again, this CFA should estimate the default measurement model though which two correlated latent factors generate a set
of 20 observed indicators.

------------------------------------------------------------------------------------------------------------------------

###

Sketch a path diagram of the CFA model described above.

- Use pencil and paper (or the equivalent); don't generate the diagram programmatically.

------------------------------------------------------------------------------------------------------------------------

### {#cfaSyntax2}

Define the `lavaan` model syntax for a two-factor model of the *Warmth* and *Dominance* items.

- Save this syntax as an object in your environment.

<details>
  <summary>Click to show code</summary>

```{r}
mod2 <- '
warmth    =~ A1 + A2 + A3 + A4 + A5 + A6 + A7 + A8 + A9 + A10
dominance =~ D1 + D2 + D3 + D4 + D5 + D6 + D7 + D8 + D9 + D10
'
```

</details>

------------------------------------------------------------------------------------------------------------------------

### {#cfaEstimation2}

Estimate the two-factor model you specified in \@ref(cfaSyntax2), and summarize the results.

- Identify the model by fixing the first factor loading of each construct to 1.
- Do not estimate any mean structure.

<details>
  <summary>Click to show code</summary>

```{r}
## Estimate the two factor model:
outH2.1 <- cfa(mod2, data = cattell)

## Summarize the results:
summary(outH2.1)
```

</details>

------------------------------------------------------------------------------------------------------------------------

###

Report the following parameter estimates from the model you estimated in \@ref(cfaEstimation2).

1. The `warmth` $\rightarrow$ `A3` factor loading
1. The `dominance` $\rightarrow$ `D2` factor loading
1. The latent covariance between *warmth* and *dominance*

<details>
  <summary>Click to show answer</summary>
  
```{r, echo = FALSE}
lambda <- inspect(outH2.1, "est")$lambda
lam3_1 <- lambda[3, 1] |> as.numeric() |> round(3)
lam12_2 <- lambda[12, 2] |> as.numeric() |> round(3)
psi2_1 <- inspect(outH2.1, "est")$psi[2, 1] |> as.numeric() |> round(3)
```

1. $\lambda_{3,1} = `r lam3_1`$
1. $\lambda_{12,2} = `r lam12_2`$
1. $\psi_{2,1} = `r psi2_1`$

</details>

------------------------------------------------------------------------------------------------------------------------

###

Based on the CFA you estimated in \@ref(cfaEstimation2), can we infer a significant linear association between *Warmth*
and *Dominance*?

- Use the statistics from your estimated model to justify your conclusion.

<details>
  <summary>Click to show answer</summary>
  
```{r, echo = FALSE}
tmp <- parameterEstimates(outH2.1) |> filter(lhs == "warmth", op == "~~", rhs == "dominance")
se <- round(tmp$se, 2)
z <- round(tmp$z, 2)
pValue <- prettyPValue(tmp$p)
```

Yes, warmth and dominance are significantly correlated
($\psi_{2,1} = `r psi2_1`$, $SE = `r se`$, $z = `r z`$, $`r pValue`$).

</details>

------------------------------------------------------------------------------------------------------------------------

###

Use the `semPaths()` function to draw a path diagram of your estimated CFA.

- Include the parameter estimates in the diagram.

<details>
  <summary>Click to show code</summary>

```{r}
semPaths(outH2.1, whatLabels = "estimates")
```

</details>

------------------------------------------------------------------------------------------------------------------------

### {#cfaEstimation3}

Estimate a modified version of the two-factor CFA from \@ref(cfaEstimation2), and summarize the results.

- Estimate the mean structure.
- Identify the model with the fixed factor method (i.e., standardize the latent variables).

<details>
  <summary>Click to show code</summary>

```{r}
outH2.2 <- cfa(mod2, data = cattell, meanstructure = TRUE, std.lv = TRUE)
summary(outH2.2)
```

</details>

------------------------------------------------------------------------------------------------------------------------

###

Report the following parameter estimates from the model you estimated in \@ref(cfaEstimation3).

1. The item intercept for `A1`
1. The item intercept for `D5`
1. The mean of the *Warmth* factor

<details>
  <summary>Click to show answer</summary>
  
```{r, echo = FALSE}
tau <- inspect(outH2.2, "est")$nu[c("A1", "D5"), ] |> as.numeric() |> round(2)
```

1. $\tau_1 = `r tau[1]`$
1. $\tau_{15} = `r tau[2]`$
1. This is a bit of a trick question. We fixed the latent means to zero for model identification, so we don't have an
estimated latent mean for the warmth factor.

</details>

------------------------------------------------------------------------------------------------------------------------

###

Based on the CFA you estimated in \@ref(cfaEstimation3), how strong is the association between *Warmth* and *Dominance*?

- Use the statistics from your estimated model to justify your conclusion.

<details>
  <summary>Click to show answer</summary>
  
```{r, echo = FALSE}
psi <- inspect(outH2.2, "est")$psi[2, 1] |> as.numeric() |> round(3)
```

We set the scale by standardizing the latent factors, so we can interpret the latent covariance as a correlation. Hence,
the estimated latent covariance of $\psi_{2,1} = `r psi`$ suggests a weak linear association between *Warmth* and
*Dominance*.

</details>

------------------------------------------------------------------------------------------------------------------------

###

Use the `semPaths()` function to draw a path diagram of the CFA you estimated in \@ref(cfaEstimation3).

- Include the parameter estimates in the diagram.

<details>
  <summary>Click to show code</summary>

```{r}
semPaths(outH2.2, whatLabels = "estimates")
```

</details>

------------------------------------------------------------------------------------------------------------------------

End of At-Home Exercises

------------------------------------------------------------------------------------------------------------------------

[ipip]: https://ipip.ori.org/
[cattell_wiki]: https://en.wikipedia.org/wiki/16PF_Questionnaire

