## In-Class Exercises

```{r, include = FALSE}
knitr::opts_chunk$set(error = TRUE)
options(show.signif.stars = FALSE)
```

In these exercises, we'll continue the factor analysis of the ESS data that we started in the At-Home Exercises.

------------------------------------------------------------------------------------------------------------------------

### Basic CFA

------------------------------------------------------------------------------------------------------------------------

####

Load the ESS data.

- The relevant data are contained in the [*ess_round1.rds*][ess_data] file.

<details>
  <summary>Click to show code</summary>
  
```{r}
dataDir <- "data"
ess <- readRDS(here::here(dataDir, "ess_round1.rds"))
```

</details>

------------------------------------------------------------------------------------------------------------------------

We are going to extend our existing three-factor CFA by adding the five-factor representation of the *Attitudes toward
Immigration* items suggested by Kestilä (2006). The hypothesized factor-to-item mapping is given below.

```{r, echo = FALSE, include = FALSE}
xxx <- readRDS(here::here(dataDir, "ess_variable_descriptions.rds"))
```

**Trust in Institutions**

- `trstlgl`: `r xxx$trstlgl`
- `trstplc`: `r xxx$trstplc`
- `trstun`: `r xxx$trstun`
- `trustep`: `r xxx$trstep`
- `trustprl`: `r xxx$trstprl`

**Satisfaction with the Political Situation**

- `stfhlth`: `r xxx$stfhlth`
- `stfedu`: `r xxx$stfedu`
- `stfeco`: `r xxx$stfeco`
- `stfgov`: `r xxx$stfgov`
- `stfdem`: `r xxx$stfdem`

**Trust in Politicians**

- `pltinvt`: `r xxx$pltinvt`
- `pltcare`: `r xxx$pltcare`
- `trstplt`: `r xxx$trstplt`

**Immigration Policy**

- `imrcntr`: `r xxx$imrcntr`
- `eimrcnt`: `r xxx$eimrcnt`
- `eimpcnt`: `r xxx$eimpcnt`
- `imsmetn`: `r xxx$imsmetn`
- `impcntr`: `r xxx$impcntr`
- `imdfetn`: `r xxx$imdfetn`

**Social Threat**

- `imbgeco`: `r xxx$imbgeco`
- `imbleco`: `r xxx$imbleco`
- `imwbcnt`: `r xxx$imwbcnt`
- `imwbcrm`: `r xxx$imwbcrm`
- `imtcjob`: `r xxx$imtcjob`
- `imueclt`: `r xxx$imueclt`

**Refugee Policy**

- `gvrfgap`: `r xxx$gvrfgap`
- `imrsprc`: `r xxx$imrsprc`
- `rfgbfml`: `r xxx$rfgbfml`
- `rfggvfn`: `r xxx$rfggvfn`
- `rfgawrk`: `r xxx$rfgawrk`
- `rfgfrpc`: `r xxx$rfgfrpc`
- `shrrfg`: `r xxx$shrrfg`

**Cultural Threat**

- `qfimchr`: `r xxx$qfimchr`
- `qfimwht`: `r xxx$qfimwht`
- `pplstrd`: `r xxx$pplstrd`
- `vrtrlg`: `r xxx$vrtrlg`

**Economic Threat**

- `imwgdwn`: `r xxx$imwgdwn`
- `imhecop`: `r xxx$imhecop`

------------------------------------------------------------------------------------------------------------------------

####

Calculate the following quantities for the CFA described above.

1. The number of model parameters
1. The pieces of available information provided by the data
1. The degrees of freedom after sufficiently identifying the model

Assume:

- Simple structure with no cross-loadings
- Enough constraints to locally identify each construct
- No mean structure

<details>
  <summary>Click to show answer</summary>

Our model will have 38 observed indicators and 8 latent constructs, so we get the following values.

**Model Parameters**

- 8 (8 - 1) / 2 = 28 latent covariances
- 8 latent variances
- 38 factor loadings
- 38 residual variances

Total = 28 + 8 + 2 $\times$ 38 = 112

**Available Information**

38 (38 + 1) / 2 = 741

**Degrees of Freedom**

To identify the model, we need to constrain one parameter in each construct's sub-model. We also need to constrain
another parameter in the sub-model for *Economic Threat*, because that construct has only two indicators. So, we need to
constrain 9 parameters.

df = 741 - 112 + 9 = 638

</details>

------------------------------------------------------------------------------------------------------------------------

#### {#cfaSyntaxClass5.1}

Define the `lavaan` model syntax for the CFA described above.

- Enforce a simple structure: do not allow any cross-loadings.
- Covary all latent factors.
- Do not specify any mean structure.
- Save this model syntax as an object in your environment.

<details>
  <summary>Click to show code</summary>

```{r}
mod_trust_att <- '
## Trust in Institutions:
inst =~ trstlgl + trstplc + trstun + trstep + trstprl

## Satisfaction with the Political Situation:
sat =~ stfhlth + stfedu  + stfeco + stfgov + stfdem

## Trust in Politicians:
pol =~ pltinvt + pltcare + trstplt

## Immigration Policy:
immi =~ imrcntr + eimrcnt + eimpcnt + imsmetn + impcntr + imdfetn 

## Social Threat:
social =~ imbgeco + imbleco + imwbcnt + imwbcrm + imtcjob + imueclt

## Refugee Policy:
refugee =~ gvrfgap + imrsprc + rfgbfml + rfggvfn + rfgawrk + rfgfrpc + shrrfg

## Cultural Threat:
cult =~ qfimchr + qfimwht + pplstrd + vrtrlg

## Economic Threat:
econ =~ l*imwgdwn + l*imhecop 
'
```

*Note:* We equate the factor loadings for the *Economic Threat* factor because that construct is under-identified
without the extra constraints.

</details>

------------------------------------------------------------------------------------------------------------------------

#### {#estimateCfaTrustAtt}

Estimate the CFA model you defined above, and summarize the results.

- Use the `lavaan::cfa()` function to estimate the model.
- Use the fixed-factor method of identification
- Use the default settings for the `cfa()` function.
- Request the model fit statistics and standardized parameter estimates by specifying `fit.measures = TRUE` and
`standardized = TRUE` in `summary()`.


<details>
  <summary>Click to show code</summary>
  
```{r}
## Load the lavaan package:
library(lavaan)

## Estimate the CFA model:
out_trust_att <- cfa(mod_trust_att, data = ess, std.lv = TRUE)

## Summarize the fitted model:
summary(out_trust_att, fit.measures = TRUE, standardized = TRUE)
```
 
</details>

------------------------------------------------------------------------------------------------------------------------

####

Evaluate the model you estimated in \@ref(estimateCfaTrustAtt). Do you think this measurement model is a reasonable
representation of the data?

<details>
  <summary>Click for explanation</summary>

```{r, echo = FALSE, include = FALSE}
fit <- fitMeasures(out_trust_att) |> as.list()

x2 <- fit$chisq |> round(2) |> format(scientific = FALSE)
df <- fit$df |> round()
p  <- prettyPValue(fit$pvalue)

rmsea <- fit$rmsea |> round(3)
cfi   <- fit$cfi |> round(3)
srmr  <- fit$srmr |> round(3)

n <- fit$ntotal |> format(scientific = FALSE)
```

**Model Fit**

The model fit is ambiguous
$(\chi^2 = `r x2`,$
$\textit{df} = `r df`,$
$`r p`,$
$\textrm{RMSEA} = `r rmsea`,$
$\textrm{CFI} = `r cfi`,$
$\textrm{SRMR} = `r srmr`)$.

- The SRMR and RMSEA look good.
- The CFI indicates poor fit.
- The $\chi^2$ is highly significant, but we don't care.

Two out of our three preferred approximate fit indices suggest a good model, but we need to be cautious. The RMSEA can
be overly optimistic when computed from large models with many degrees of freedom. So, we have a bit of a tie:

- One good-looking index (SRMR)
- One bad-looking index (CFI)
- One good-looking, but not entirely trustworthy, index (RMSEA)

**Parameter Estimates**

The standardized factor loadings and residual variances seem fine.

```{r}
est <- lavInspect(out_trust_att, "standardized")
cbind(est$lambda, Resid_Var = diag(est$theta)) |>
  round(3) |> 
  Matrix::Matrix(sparse = TRUE)
```

The latent correlations look sensible.

```{r}
round(est$psi, 3)
```

The balance of evidence suggests that this model is probably useful, but we're definitely in a gray area. In practice,
we would need to consider the context of our analysis. For example, we should ask ourselves questions like the following.

*How noisy are the data?*

Do we have high-quality data that should support an unambiguously good model? If so, these results don't stack up. On
the other hand, we might not have high expectations because we know that our data are very noisy. If we're just using
this analysis to provide some vague guidance toward next steps in our research line, we may be fine with this model.

*What are the stakes of "getting it wrong"?*

We don't want to use this quality of result to make mission-critical inference. For example, if our results will
influence real-world policy or impact the design of subsequent projects in our research line, this model isn't good
enough.

*What are the standards in our field?*

If we're working in a field where noisy data and weak theory are unavoidable, then this model is probably good enough to
provide some meaningful insight. That's not to say that we should promulgate questionable research practices. If good
measurement and rigorous theory are possible, we should shoot for those more lofty objectives. Frequently, however, we
really don't have any hope of a good fitting model. In such situations, we should make the most of the data we can get.

</details>

------------------------------------------------------------------------------------------------------------------------

When we fail to support the hypothesized measurement model, the confirmatory phase of our analysis is over. At this
point, we've essentially rejected our hypothesized measurement structure, and that's the conclusion of our analysis. We
don't have to throw up our hands in despair, however. We can still contribute something useful by modifying the
theoretical measurement model through an exploratory, data-driven, post-hoc analysis.

We'll give that a shot below.

------------------------------------------------------------------------------------------------------------------------

### Post-Hoc Model Modifications

------------------------------------------------------------------------------------------------------------------------

We should avoid making any post-hoc model modifications until we don't have any other choice (i.e., our hypothesized
model is clearly rejected). Unfortunately, such cases arise frequently in real-world data analysis. So, we want to
make the most of these bad situations. 

#### Modification Indices {-}

If we're going to modify our hypothesized model structure, the \emph{modification indices} can help guide us.

Modification indices estimate the improvement in model fit that we'd get from freeing parameters that are fixed in the
current model. If a parameter has a large modification index, we expect substantially better fit for a model that freely
estimates that parameter. Hence, we can use the modification indices to help us choose which parameter estimates to add
when modifying our hypothesized model.

The following code will compute the modification indices for the three-factor model of *Trust in Politics* from the
At-Home Exercises.

```{r, paged.print = FALSE}
modificationIndices(out_3f, sort. = TRUE)
```

As you can see, the `modificationInices()` function will usually produce a lot of output, but we don't need all that
stuff. We only care about the largest modification indices, so we only need to print the first few rows.

```{r, paged.print = FALSE}
modificationIndices(out_3f, sort. = TRUE) |> head()
```

Alternatively, we can request only those modification indices that exceed some minimum threshold. We can use 10\% of the
model $\chi^2$ as a useful rule-of-thumb.

```{r, paged.print = FALSE}
thresh <- 0.1 * fitMeasures(out_3f, "chisq")
modificationIndices(out_3f, sort. = TRUE, minimum.value = thresh)
```

```{r, echo = FALSE, include = FALSE}
mi <- modificationIndices(out_3f, sort. = TRUE)[1, , drop = FALSE]
v1 <- mi$lhs
v2 <- mi$rhs
val <- mi$mi |> round(2)
text <- ifelse(mi$op == "~~", 
  paste0("covariance between `", v1, "` and `", v2, "`"),
  paste0("factor loading from `", v1, "` to `", v2, "`")
)
```

So, we can infer that freeing the `r text` would reduce the model $\chi^2$ by `r val`, for example.

#### Using Modification Indices {-}

To keep our modifications in check, we should follow the guidelines below when using modification indices to help choose
which parameters to free.

- Only consider parameters with modification indices larger than 10\% of the model $\chi^2$.
- Only consider parameters that make obvious theoretical sense.
- Only consider parameters that you believe to be part of the true model.

The final guideline is particularly important. The goal of post-hoc model modifications is not getting the model to fit
our data: that ship has already sailed. We're proposing changes to the theoretical model for the benefit of future
research on our topic. That's a heavy task. We can only make a justified post-hoc model modification if we can honestly
agree with the following statements.

- My a priori hypothesized model was wrong.
- These modifications are part of the true population model that I'm trying to estimate.
- All future analyses should use this modified version of the model as the a priori hypothesized model.

To give you a sense of the gravity of these decisions, consider the following humble-brag. I've been doing this type of
modeling in some professional capacity for more than 15 years, and I can't recall a single case where I've been able to
make a truly justifiable post-hoc model modification. That's not to say that I haven't modified my fair share of
models, but this is one of those "do as I say, not as I do" type situations.

------------------------------------------------------------------------------------------------------------------------

####

Calculate the modification indices for the model you estimated in \@ref(estimateCfaTrustAtt).

- Which estimate would most improve the model fit?
- How much would freeing the above estimate decrease the model $\chi^2$.
- Can you see an obvious reason why this parameter estimate would improve the model?
   - If so, explain the reason.

<details>
  <summary>Click to show code</summary>

```{r, paged.print = FALSE}
thresh <- 0.1 * fitMeasures(out_trust_att, "chisq")
modificationIndices(out_trust_att, sort. = TRUE, minimum.value = thresh)
```

  <details>
    <summary>Click for explanation</summary>

```{r, echo = FALSE, include = FALSE}
mi <- modificationIndices(out_trust_att, sort. = TRUE)[1, , drop = FALSE]
v1 <- mi$lhs
v2 <- mi$rhs
val <- mi$mi |> round(2)
text <- ifelse(mi$op == "~~", 
  paste0("covariance between `", v1, "` and `", v2, "`"),
  paste0("factor loading from `", v1, "` to `", v2, "`")
)
```

- Freeing the `r text` would produce the largest improvement in model fit.
- After freeing this parameter, we'd expect the model $\chi^2$ to decrease by `r val`.
- Yes. Both of these items ask about attitudes towards immigrants from rich countries. None of the other items
  explicitly address immigration from rich countries.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

#### {#estimateCfaTrustAtt2}

Estimate a modified version of the CFA from \@ref(estimateCfaTrustAtt).

- Modify the model to estimate the parameter associated with the largest modification index that you found above.

<details>
  <summary>Click to show code</summary>

```{r}
## Add the residual covariance to the model syntax:
mod_trust_att2 <- paste(mod_trust_att, "imrcntr ~~ eimrcnt", sep = "\n")

## Estimate and summarize the modified model:
out_trust_att2 <- cfa(mod_trust_att2, data = ess, std.lv = TRUE)
summary(out_trust_att2, fit.measures = TRUE, standardized = TRUE)
```

</details>

------------------------------------------------------------------------------------------------------------------------

####

Calculate the modification indices for the model you estimated in \@ref(estimateCfaTrustAtt2).

- Which estimate would most improve the model fit?
- How much would freeing the above estimate decrease the model $\chi^2$.
- Can you see an obvious reason why this parameter estimate would improve the model?
   - If so, explain the reason.

<details>
  <summary>Click to show code</summary>

```{r, paged.print = FALSE}
thresh <- 0.1 * fitMeasures(out_trust_att2, "chisq")
modificationIndices(out_trust_att2, sort. = TRUE, minimum.value = thresh)
```

  <details>
    <summary>Click for explanation</summary>

```{r, echo = FALSE, include = FALSE}
mi <- modificationIndices(out_trust_att, sort. = TRUE)[1, , drop = FALSE]
v1 <- mi$lhs
v2 <- mi$rhs
val <- mi$mi |> round(2)
text <- ifelse(mi$op == "~~", 
  paste0("covariance between `", v1, "` and `", v2, "`"),
  paste0("factor loading from `", v1, "` to `", v2, "`")
)
```

- Freeing the `r text` would produce the largest improvement in model fit.
- After freeing this parameter, we'd expect the model $\chi^2$ to decrease by `r val`.
- Yes. Both of these items ask about trust in parts of the legal system. None of the other questions address trust in
courts or police.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

#### {#estimateCfaTrustAtt3}

Estimate a modified version of the CFA from \@ref(estimateCfaTrustAtt).

- Modify the model to estimate the parameter associated with the largest modification index that you found above.

<details>
  <summary>Click to show code</summary>

```{r}
## Add the residual covariance to the model syntax:
mod_trust_att3 <- paste(mod_trust_att2, "trstlgl ~~ trstplc", sep = "\n")

## Estimate and summarize the modified model:
out_trust_att3 <- cfa(mod_trust_att3, data = ess, std.lv = TRUE)
summary(out_trust_att3, fit.measures = TRUE, standardized = TRUE)
```

</details>

------------------------------------------------------------------------------------------------------------------------

### Model Comparisons

------------------------------------------------------------------------------------------------------------------------

For the At-Home Exercises, you qualitatively compared the fit of a three-factor *Trust in Politics* model to the fit of
a one-factor version. This type of qualitative comparison is fine, but we'd like to have an actual statistical test for
these fit differences. As it happens, we have just such a test: a nested model $\Delta \chi^2$ test (AKA, chi-squared
difference test, change in chi-squared test, likelihood ratio test).

In the coming weeks, we'll cover nested models and tests thereof, but it will be useful to start thinking about these
concepts now. Two models are said to be *nested* if you can define one model by placing constraints on the other model.

By way of example, consider the following two CFA models.

```{r, echo = FALSE}
library(semPlot)
library(semptools)

hsMod <- '
visual  =~ x1 + x2 + x3
textual =~ x4 + x5 + x6
'

hsFit <- cfa(hsMod, data = HolzingerSwineford1939) 

semPaths(hsFit,
         layout = "tree2",
         sizeMan = 10,
         sizeMan2 = 8,
         sizeLat = 18,
         curve = 3,
         esize = 2,
         asize = 4,
         mar = c(6, 2, 10, 2),
         DoNotPlot = TRUE,
         nCharNodes = 0) %>%
rotate_resid(c(visual = -45, textual = 45)) %>%
plot()

hsMod2 <- '
visual  =~ x1 + x2 + x3
textual =~ x4 + x5 + x6

visual ~~ 0 * textual
'

cfa(hsMod2, data = HolzingerSwineford1939) %>%
  semPaths(layout = "tree2",
           sizeMan = 10,
           sizeMan2 = 8,
           sizeLat = 18,
           curve = 3,
           esize = 2,
           asize = 4,
           mar = c(6, 2, 10, 2),
           DoNotPlot = TRUE,
           nCharNodes = 0,
           exoCov = FALSE) %>%
rotate_resid(c(visual = -45, textual = 45)) %>%
plot()
```

The second model is nested within the first model, because we can define the second model by fixing the latent
covariance to zero in the first model.

Notice that the data contain $6(6 + 1) / 2 = 21$ unique pieces of information. The first model estimates
`r (p = fitMeasures(hsFit, "npar"))` parameters, and the second model estimates `r p - 1` parameters. Hence the first
model has `r (df1 = 21 - p)` degrees of freedom, and the second model has `r df1 + 1` degrees of freedom.

In general, the following must hold whenever *Model B* is nested within *Model A*.

- Model B will have *fewer* estimated parameters than Model A.
- Model B will have *more* degrees of freedom than Model A.
- Model A will be *more* complex than model B.
- Model A will fit the data *better* than model B.

------------------------------------------------------------------------------------------------------------------------

**Saturated Model**

All models are nested within the saturated model, because the saturated model estimates all possible relations among the
variables. Regardless of what model we may be considering, we can always convert that model to a saturated model by
estimating all possible associations. Hence, all models are nested within the saturated model.

```{r, echo = FALSE}
tmp        <- rep(0, 6)
names(tmp) <- lavNames(hsFit)

lav_partable_unrestricted(hsFit) %>% 
  cfa(data = HolzingerSwineford1939) %>%
  semPaths(layout = "tree",
           sizeMan = 10,
           sizeMan2 = 8,
           curve = 3,
           esize = 2,
           asize = 4,
           DoNotPlot = TRUE) %>%
  rotate_resid(tmp) %>%
  plot()
```

------------------------------------------------------------------------------------------------------------------------

**Baseline Model**

Similarly, the baseline model (AKA, independence model) is nested within all other models. In the baseline model, we
only estimate the variances of the observed items; all associations are constrained to zero. We can always convert our
model to the baseline model by fixing all associations to zero. Hence, the baseline model is nested within all other
models.

```{r, echo = FALSE}
lav_partable_independence(hsFit) %>% 
  cfa(data = HolzingerSwineford1939) %>%
  semPaths(layout = "tree",
           sizeMan = 10,
           sizeMan2 = 8,
           esize = 2,
           asize = 4)
```

------------------------------------------------------------------------------------------------------------------------

When two models are nested, we can use a $\Delta \chi^2$ test to check if the nested model fits *significantly worse*
than its parent model. Whenever we place constraints on the model, the fit will deteriorate, but we want to know if the
constraints we imposed to define the nested model have produced *too much* loss of fit.

We can use the `anova()` function to easily conduct $\Delta \chi^2$ tests comparing models that we've estimated with
`cfa()` or `sem()`.

------------------------------------------------------------------------------------------------------------------------

####

Use the `anova()` function to compare the three-factor model *Trust in Politics* from \@ref(estimateCfa3F) to the
one-factor model from \@ref(estimateCfa1F). 

- Explain what `Df`, `Chisq`, `Chisq diff`, `Df diff`, and `Pr(>Chisq)` mean.
- Which model is more complex?
- Which model fits better?
- What is the conclusion of the test?

<details>
  <summary>Click to show code</summary>
  
```{r}
anova(out_1f, out_3f)
```
  
  <details>
    <summary>Click for explanation</summary>
    
The `Df` column contains the degrees of freedom of each model.

- Higher df $\Rightarrow$ Less complex model

The `Chisq` column shows the $\chi^2$ statistics (AKA, likelihood ratio statistics) for each model.

- $\chi^2$ = The ratio of the likelihoods for the estimated model and the saturated model.
- Larger $\chi^2$ $\Rightarrow$ Worse fit

`Chisq diff` is the difference between the two $\chi^2$ values (i.e., $\Delta \chi^2$).

- How much better the more complex model fits the data
- Larger $\Delta \chi^2$ values indicate greater losses of fit induced by the constraints place on the nested model.

`Df diff` is the difference in the degrees of freedom between the models.

- Since both models must be estimated from the same pool of variables, this difference also represents the number of
parameters that were constrained to define the nested model.
 
`Pr(>Chisq)` is a p-value for the $\Delta \chi^2$ test.

- $H_0: \Delta \chi^2 = 0$
- $H_1: \Delta \chi^2 > 0$

The three-factor model is more complex than the one-factor model, but the extra complexity is justified. The
three-factor model fits significantly better than the one-factor model.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

####

Use a $\Delta\chi^2$ test to test if the additional parameter estimates you added in \@ref(estimateCfaTrustAtt2) and
\@ref(estimateCfaTrustAtt3) significantly improve the model fit.

- Did the model fit significantly improve?
- Is the fit of the modified model acceptable?

<details>
  <summary>Click to show code</summary>
  
```{r, paged.print = FALSE}
anova(out_trust_att, out_trust_att3)
fitMeasures(out_trust_att3)
```
  
  <details>
    <summary>Click for explanation</summary>
 
```{r, echo = FALSE, include = FALSE}
aOut <- anova(out_trust_att, out_trust_att3)
dX2  <- aOut[2, "Chisq diff"] |> round(2)
dDf  <- aOut[2, "Df diff"] |> round()
dP   <- prettyPValue(aOut[2, "Pr(>Chisq)"])

fit <- fitMeasures(out_trust_att3) |> as.list()

x2 <- fit$chisq |> round(2) |> format(scientific = FALSE)
df <- fit$df |> round()
p  <- prettyPValue(fit$pvalue)

rmsea <- fit$rmsea |> round(3)
cfi   <- fit$cfi |> round(3)
srmr  <- fit$srmr |> round(3)
```

Yes, these modifications significantly improve the model fit
$(\Delta\chi^2 = `r dX2`,$
$\textit{df} = `r dDf`,$
$`r dP`).$
In this case, the original 8-factor model is nested within the modified model. So, our $\Delta \chi^2$ test is
evaluating the improvement in fit contributed by freeing the two residual covariances. The $\Delta \chi^2$ test is
significant, so we can conclude that including the two new parameter estimates has significantly improved the model fit.

- I.e., Estimating these two residual covariances is "worth it" in the sense of balancing model fit and model complexity.

Also, the overall fit of the modified model is now more unanbiguously acceptable
$(\chi^2 = `r x2`,$
$\textit{df} = `r df`,$
$`r p`,$
$\textrm{RMSEA} = `r rmsea`,$
$\textrm{CFI} = `r cfi`,$
$\textrm{SRMR} = `r srmr`)$.

**Very Important Caveat!**

If we had found this result when testing our original model, we would be well-situated to proceed with our analysis. In
this case, however, we are no longer justified in generalizing these estimates to the population. We only arrived at
this well-fitting model by modifying our original theoretical model to better fit the data using estimates derived from
those same data to guide our model modifications.

We've conducted this post-hoc analysis to help inform future research, and this result is useful as a starting point for
future studies. Now, anyone analyzing these scales in the future could incorporate these residual covariances into their
initial theoretical model. Basically, we conduct these types of post-hoc analyses to help future researchers learn from
our mistakes.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

#### {#estimateCfaTrustAtt0}

Estimate a restricted version of the 8-factor CFA from \@ref(estimateCfaTrustAtt).

- Remove all latent covariances between the three *Trust in Politics* factors and the five *Attitudes Toward
Immigration* factors.
- Retain all other aspects of the original CFA.

Summarize the estimated model to check that everything worked.

<details>
  <summary>Click to show code</summary>

First, we'll define a syntax snippit to enforce the constraints.

```{r}
constraints <- '
inst ~~ 0*immi + 0*social + 0*refugee + 0*cult + 0*econ
sat  ~~ 0*immi + 0*social + 0*refugee + 0*cult + 0*econ
pol  ~~ 0*immi + 0*social + 0*refugee + 0*cult + 0*econ
'
```

Now, we'll add the constraint syntax to the CFA syntax.

```{r}
mod_trust_att0 <- paste(mod_trust_att, constraints, sep = "\n")
```

Finally, we'll estimate and summarize the restricted model.

```{r}
out_trust_att0 <- cfa(mod_trust_att0, data = ess, std.lv = TRUE)
summary(out_trust_att0, fit.measures = TRUE, standardized = TRUE)
```

</details>

------------------------------------------------------------------------------------------------------------------------

Below, you will conduct a $\Delta\chi^2$ test to compare the full 8-factor CFA from \@ref(estimateCfaTrustAtt) to the
restricted version you estimated in \@ref(estimateCfaTrustAtt0). All remaining questions refer to this test.

------------------------------------------------------------------------------------------------------------------------

####

1. Which model is more complex?
1. Which model has more degrees of freedom?
1. What is the difference in the degrees of freedom between the models?
1. How many degrees of freedom will we have for this test?

<details>
  <summary>Click for explanation</summary>

1. The unrestricted CFA is more complex: we estimate more parameters in that model.
1. The restricted model will have more degrees of freedom: we lose one degree of freedom for every estimated parameter.
1. $\Delta \textit{df} = 24$: We fixed 24 latent covariances to define the restricted model, so the restricted model has
   24 more degrees of freedom.
1. Our $\chi^2$ test will have $\textit{df} = 24$: The df for a $\Delta\chi^2$ is equal to the $\Delta \textit{df}$
   between the compared models.

</details>

------------------------------------------------------------------------------------------------------------------------

####

Use R to conduct the $\Delta\chi^2$ test.

- What is the conclusion of the test?
- Support your claim with appropriate statitsics.

<details>
  <summary>Click to show code</summary>
  
```{r, paged.print = FALSE}
anova(out_trust_att, out_trust_att0)
```
  
  <details>
    <summary>Click for explanation</summary>
    
```{r, echo = FALSE, include = FALSE}
aOut <- anova(out_trust_att, out_trust_att0)
x2   <- aOut[2, "Chisq diff"] |> round(2)
df   <- aOut[2, "Df diff"] |> round()
p    <- prettyPValue(aOut[2, "Pr(>Chisq)"])
```

The unrestricted CFA fits significantly better than the restricted CFA
$(\Delta\chi^2 = `r x2`,$
$\textit{df} = `r df`,$
$`r p`).$
Hence, the constraints imposed on the restricted model are untenable. The unrestricted CFA produces a significantly
better representation of the data.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

End of In-Class Exercises

------------------------------------------------------------------------------------------------------------------------
