## At-Home Exercises

```{r setupH5, include = FALSE, echo = FALSE}
source(here::here("code", "supportFunctions.R"))
```

In these exercises, you will attempt to confirm the measurement model structure proposed in:

Kestilä, E. (2006). Is there demand for radical right populism in the Finnish electorate? *Scandinavian Political
Studies 29*(3), 169--191. <https://doi.org/10.1111/j.1467-9477.2006.00148.x>

The data for this practical were collected during the first round of the European Social Survey (ESS). The ESS is a
repeated cross-sectional survey administered in 32 European countries. The first wave was collected in 2002, and two new
waves have been collected each year since. You can find more info and access the data at
<https://www.europeansocialsurvey.org>. 

The data we will analyze in this practical are contained in the file named [*ess_round1.rds*][ess_data]. This file
contains a processed subset of the ESS data. The following table describes the variables in this dataset.

```{r, results = "asis", echo = FALSE}
library(kableExtra)

dataDir <- "data"

## Read the 'ess_round1.rds' data into a data frame called 'ess': 
ess <- readRDS(here::here(dataDir, "ess_round1.rds"))

desc <- c(
  "Title of dataset",
  "ESS round",
  "Edition",
  "Production date",
  "Country",
  "Respondent's identification number",
  "Trust in the legal system",
  "Trust in the police",
  "Trust in the United Nations",
  "Trust in the European Parliament",
  "Trust in country's parliament",
  "State of health services in country nowadays",
  "State of education in country nowadays",
  "How satisfied with present state of economy in country",
  "How satisfied with the national government",
  "How satisfied with the way democracy works in country",
  "Politicians interested in votes rather than peoples opinions",
  "Politicians in general care what people like respondent think",
  "Trust in politicians",
  "Allow many/few immigrants of same race/ethnic group as majority",
  "Allow many/few immigrants of different race/ethnic group from majority",
  "Allow many/few immigrants from richer countries in Europe",
  "Allow many/few immigrants from poorer countries in Europe",
  "Allow many/few immigrants from richer countries outside Europe",
  "Allow many/few immigrants from poorer countries outside Europe",
  "Qualification for immigration: christian background",
  "Qualification for immigration: be white",
  "Average wages/salaries generally brought down by immigrants",
  "Immigrants harm economic prospects of the poor more than the rich",
  "Immigrants take jobs away in country or create new jobs",
  "Taxes and services: immigrants take out more than they put in or less",
  "Immigration bad or good for country's economy",
  "Country's cultural life undermined or enriched by immigrants",
  "Immigrants make country worse or better place to live",
  "Immigrants make country's crime problems worse or better",
  "Richer countries should be responsible for accepting people from poorer countries",
  "Better for a country if almost everyone share customs and traditions",
  "Better for a country if a variety of different religions",
  "Country has more than its fair share of people applying refugee status",
  "People applying refugee status allowed to work while cases considered",
  "Government should be generous judging applications for refugee status",
  "Most refugee applicants not in real fear of persecution own countries",
  "Financial support to refugee applicants while cases considered",
  "Granted refugees should be entitled to bring close family members",
  "Gender",
  "Year of birth",
  "Highest level of education",
  "Years of full-time education completed",
  "How interested in politics",
  "Placement on left right scale",
  "A factor version of the 'cntry' variable",
  "A factor version of the 'gndr' variable"
  )

data.frame(Variable = names(ess), Type = descriptives(ess)$type, Description = desc) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

------------------------------------------------------------------------------------------------------------------------

###

Load the *ess_round1.rds* dataset.

- Inspect the data after loading to make sure everything went well.

<details>
  <summary>Click to show code</summary>

```{r}
library(tidySEM)

dataDir <- "data"

## Read the 'ess_round1.rds' data into a data frame called 'ess': 
ess <- readRDS(here::here(dataDir, "ess_round1.rds"))

## Inspect the result:
dim(ess)
head(ess)
descriptives(ess)
```

</details>

------------------------------------------------------------------------------------------------------------------------

```{r, echo = FALSE, include = FALSE}
names(desc) <- colnames(ess)
xxx <- as.list(desc)
```

Kestilä, E. (2006) used principal components analysis to estimate the measurement model underlying thirteen ESS
variables related to *Trust in Politics*. Based on the results of this PCA, Kestilä recommended the following
three-factor structure.

**Trust in Institutions**

- `trstlgl`: `r xxx$trstlgl`
- `trstplc`: `r xxx$trstplc`
- `trstun`: `r xxx$trstun`
- `trustep`: `r xxx$trstep`
- `trustprl`: `r xxx$trstprl`

**Satisfaction with the Political Situation**

- `stfhlth`: `r xxx$stfhlth`
- `stfedu`: `r xxx$stfedu`
- `stfeco`: `r xxx$stfeco`
- `stfgov`: `r xxx$stfgov`
- `stfdem`: `r xxx$stfdem`

**Trust in Politicians**

- `pltinvt`: `r xxx$pltinvt`
- `pltcare`: `r xxx$pltcare`
- `trstplt`: `r xxx$trstplt`

------------------------------------------------------------------------------------------------------------------------

### {#cfaSyntax}

Define the `lavaan` model syntax for the three-factor CFA described above.

- Covary the three latent factors.
- Do not specify any mean structure.
- Save this model syntax as an object in your environment.

<details>
  <summary>Click to show code</summary>

```{r}
mod_3f <- '
institutions =~ trstlgl + trstplc + trstun + trstep + trstprl
satisfaction =~ stfhlth + stfedu  + stfeco + stfgov + stfdem
politicians  =~ pltinvt + pltcare + trstplt
'
```

</details>

------------------------------------------------------------------------------------------------------------------------

###

Estimate the CFA model you defined above, summarize the results, and evaluate the model fit.

- Use the `lavaan::cfa()` function to estimate the model.
- Use the fixed-factor method to identify the model.
- Otherwise, use the default settings for the `cfa()` function.
- Request the model fit indices by supplying the `fitMeasures = TRUE` argument to `summary()`.
- Request the standardized parameter estimates by supplying the `standardized = TRUE` argument to `summary()`.

Does the model fit the data well?

<details>
  <summary>Click the code</summary>

```{r}
## Load the lavaan package:
library(lavaan)

## Estimate the CFA model:
out_3f <- cfa(mod_3f, data = ess, std.lv = TRUE)

## Summarize the fitted model:
summary(out_3f, fit.measures = TRUE, standardized = TRUE)
```

<details>
  <summary>Click for explanation</summary>

```{r, echo = FALSE, include = FALSE}
fit <- fitMeasures(out_3f) |> as.list()

x2 <- fit$chisq |> round(2) |> format(scientific = FALSE)
df <- fit$df |> round()
p  <- prettyPValue(fit$pvalue)

rmsea <- fit$rmsea |> round(3)
cfi   <- fit$cfi |> round(3)
srmr  <- fit$srmr |> round(3)

n <- fit$ntotal |> format(scientific = FALSE)
```

No, the model does not seem to fit the data well
$(\chi^2 = `r x2`,$
$\textit{df} = `r df`,$
$`r p`,$
$\textrm{RMSEA} = `r rmsea`,$
$\textrm{CFI} = `r cfi`,$
$\textrm{SRMR} = `r srmr`)$.

- The SRMR looks good, but one good looking fit statistic is not enough.
- The RMSEA and CFI both indicate poor fit.
- The $\chi^2$ is highly significant, but we don't care (especially when our sample size is $N = `r n`$).

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

###

Consider the factor loadings for the *Trust in Politicians* factor.

- Do the loadings seem sensible?
- Do you notice any discrepancy between the raw and standardized estimates of these loadings?
- How would you explain this discrepancy (assuming you see one)?

<details>
  <summary>Click for explanation</summary>

```{r}
data.frame(
  raw = lavInspect(out_3f, "estimates")$lambda[11:13, 3],
  std = lavInspect(out_3f, "standardized")$lambda[11:13, 3]
  )
```

Hmm...something seem strange, but we need to be careful when evaluating these factor loadings.

- The raw loadings seem to suggest some pretty severe heterogeneity among the loadings of the *Trust in Politicians* factor.
- The standardized estimates don't look too bad.

The only difference between these two versions of the loadings is the scale of the indicator variables.

- The indicators for the standardized loadings all have the same unit variance.
- The indicators for the raw loadings have their original scale.

Let's check the scale of the raw indicator variables.

```{r}
ess |>
  select(lavNames(out_3f)) |>
  tidySEM::descriptives() |> 
  select(name, mean, sd, min, max)
```

Well, there's your problem! All the variables have an eleven-point scale except `pltinvt` and `pltcare`: they only have
five-point scales. So, the factor loadings for `pltinvt` and `pltcare` should be smaller than the factor loading for
`trstplt`. One unit of change in `pltinvt` or `pltcare` implies a large effect than one unit of change in `trstplt`
would.

</details>

------------------------------------------------------------------------------------------------------------------------

Now, we will consider an alternative factor structure for the *Trust in Politics* CFA. We'll go extremely simple by
estimating a one-factor model wherein all *Trust* items are explained by a single latent variable.

------------------------------------------------------------------------------------------------------------------------

###

Define the `lavaan` model syntax for a one-factor model of the *Trust* items.

- Save this syntax as an object in your environment.

<details>
  <summary>Click to show code</summary>

```{r}
mod_1f <- '
political_trust =~ 
  trstlgl +
  trstplc +
  trstun +
  trstep +
  trstprl +
  stfhlth +
  stfedu  +
  stfeco +
  stfgov +
  stfdem +
  pltinvt +
  pltcare +
  trstplt
'
```

</details>

------------------------------------------------------------------------------------------------------------------------

###

Estimate the one-factor model, and summarize the results.

- Identify the scale using the fixed factor method.
- Don't estimate any mean structure.

Does this model appear to fit better or worse than the three-factor model?

<details>
  <summary>Click to show code</summary>

```{r}
## Estimate the one factor model:
out_1f <- cfa(mod_1f, data = ess, std.lv = TRUE)

## Summarize the results:
summary(out_1f)

## Compare fit statistics:
data.frame(three_factor = fitMeasures(out_3f), one_factor = fitMeasures(out_1f)) |>
  format(scientific = FALSE, digits = 3, drop0trailing = TRUE)
```

<details>
  <summary>Click for explanation</summary>

The one-factor model definitely seems to fit worse than the three-factor model.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

End of At-Home Exercises

------------------------------------------------------------------------------------------------------------------------

[ess_data]: https://surfdrive.surf.nl/files/index.php/s/an0GVjaMiZYA73u 
