# Full SEM

```{r week7_global_options, include = FALSE}
library(knitr)
thisWeek <- "week7"
```

This week, we will focus on integrating all of the disparate methods we've covered so far into full-fledged *structural
equation models*.

**Homework before the lecture**

1. Watch the [Lecture Recording](lecture-5.html) for this week.
1. Complete the [Reading](reading-5.html) for this week, and answer the associated reading questions.

**Homework before the practical**

Complete the [At-Home Exercises](at-home-exercises-5.html). 

**Practical content**

During the practical you will work on the [In-Class Exercises](in-class-exercises-5.html).

```{r, echo = FALSE, message = FALSE, warning = FALSE, error = FALSE, results = "asis"}
knit_child(here::here("sections", thisWeek, "lecture.Rmd"), quiet = TRUE) |> cat(sep = "\n")
# knit_child(here::here("sections", "lecture-placeholder.Rmd"), quiet = TRUE) |> cat(sep = "\n")

knit_child(here::here("sections", thisWeek, "reading.Rmd"), quiet = TRUE) |> cat(sep = "\n")
# knit_child(here::here("sections", "reading-placeholder.Rmd"), quiet = TRUE) |> cat(sep = "\n")

knit_child(here::here("sections", thisWeek, "home.Rmd"), quiet = TRUE) |> cat(sep = "\n")
# knit_child(here::here("sections", "home-placeholder.Rmd"), quiet = TRUE) |> cat(sep = "\n")

knit_child(here::here("sections", thisWeek, "class.Rmd"), quiet = TRUE) |> cat(sep = "\n")
# knit_child(here::here("sections", "class-placeholder.Rmd"), quiet = TRUE) |> cat(sep = "\n")
```
