## Procedures {-}

---

### Formatting {- #formatting}

Your assignment submissions must follow certain formatting guidelines. Failing to correctly format your submissions can
impact your grade.

#### Report {-}

You must submit your Assignment 1 report in PDF format. Each report must include your name

Your report should include appendices containing the lavaan syntax used to estimate each model include in the report.
Include only the lavaan syntax string and the call to the estimation function used to fit the model (e.g., `cfa()`,
`sem()`, `lavaan()`). For example, the following snippets would be appropriate.

```{r, eval = FALSE}
cfaMod <- '
f1 =~ a1 + a2 + a3
f2 =~ b1 + b2 + b3
'

cfaOut <- cfa(cfaMod, data = dat1, std.lv = TRUE)
```

```{r, eval = FALSE}
semMod <- '
f1 =~ a1 + a2 + a3
f2 =~ b1 + b2 + b3

f1 ~ f2 + x
'

semOut <- sem(semMod, data = dat1, std.lv = TRUE)
```

The snippet below, however, contains additional unnecessary commands related to data ingest/processing and summarizing
the model.

```{r, eval = FALSE}
dat0 <- readRDS("../data/dataset.rds")
dat1 <- rename(dat0, x = foo, y = bar, z = baz)

pathMod <- 'foo ~ bar + baz'
pathOut <- sem(pathOut, data = dat1)

summary(pathOut)
```

#### Syntax File {- #analysis_script}

You must also submit an executable script containing the code used for your analyses.

- If you used [**KnitR**][knitr] to embed your analysis code in your report, you may submit the source code that
generates your report (i.e., an RMD [R Markdown], QMD [Quarto], or RNW [LaTeX] file).
- Otherwise, submit a standard R script.

Clean the script before submitting.

- Remove extraneous/redundant code and comments.
   - Only include the code for analyses that appear in your report.
   - Include code for data processing/clean/wrangling.
- Order the commands to match your final analysis.
   - E.g., Data ingest $\rightarrow$ Data processing $\rightarrow$ Descriptive analysis $\rightarrow$ Modeling
   $\rightarrow$ Processing results $\rightarrow$ Post-hoc analysis
   - In theory, I should be able to source your script (i.e., run the entire script in "batch mode") to reproduce your
   results.

---

### Length {-}

You may use as many words as necessary to adequately explain yourself, but concision and parsimony are encouraged. Note
that the assignments are **not** intended to be full-blown research papers! The focus should be on the definition of
your model, how this model relates to theory (introduction), and what you have learned from your estimated model
(discussion).

You should be able to get the job done in fewer than 10 pages of text (excluding title page, figures, appendices, and
references).

---

### Submission {-}

You will submit your A1 report through Brightspace.

- You can find a submission portal in the "Assignments" section of the Brightspace page.
- If something goes wrong with the submission, or you notice a mistake (before the deadline) that you want to correct,
you may upload a new version of your report. 
   - We will grade the final submitted version.
- The submissions will be screened with *TurnItIn* to check for plagiarism.

---

### Rules {-}

#### Resources {-}

For both assignments, you may use any reference materials you like, including:

- All course materials
- The course GitBook
- Additional books and papers
- The internet

#### Collaboration {-}

You will complete the first assignment individually. 

- For this assignment, you may not work with anyone else.

You will complete the second assignment in groups.

- Although you will work in groups, your group may not work together with other groups.

For both assignments, you are obligated to submit original work (i.e., work conducted for this course by you or your
group). 

- Submitting an assignment that violates this condition constitutes fraud.
- Such cases of fraud will be addressed according to the 
[University's standard policy](https://students.uu.nl/en/practical-information/policies-and-procedures/fraud-and-plagiarism).

#### Academic integrity {-}

Hopefully, you also feel a moral obligation to obey the rules. For this course, we have implemented an examination that
allows you to showcase what you have learned in a more realistic way than a written exam would allow. 

- This assessment format spares you the stress of long exams and the attendant studying/cramming. 
- The assignments will also help you assess your ability to independently analyse data, which is important to know for
your future courses and/or career.

However, this format also assumes that you complete the assignments in good faith. So, I simply ask that you hold up
your end of the bargain, and submit your original work to show us what you've learned.

#### GenAI disclosure statement {-}

Naturally, all of the [course policies on GenAI](#ai_policy) also apply to the data analysis assignments. In addition to
these general policies, you must include a GenAI disclosure statement in each of your assignments.

You are free to use AI tools like ChatGPT, Claude, or CoPilot, but you must cite the tools you used in the same way that
you would cite any other software. As with citing any other source, properly citing AI tools both attributes the
appropriate credit to the tool’s developers and helps you clearly and accurately describe your own work. In addition to
providing a valid citation, you must also explain how you used the AI tool. Specifically, did you use AI to:

1. Create new content (e.g., use ChatGPT to draft some parts of the text)
1. Modify content that you created (e.g., use Claude to debug your R code)

Provide this information in a separate *GenAI Disclosure Statement* section at the end of your report.

#### Strict stuff {-}

By submitting your assignments (both group and individual), you confirm the following:

1. You have completed the assignment yourself (or with your group)
1. You are submitting work that you have written yourself (or with your group)
1. You are using your own UU credentials to submit the assignment
1. You have not had outside help that violates the conditions delineated above while completing the assignment

All assignments will be submitted via TurnItIn in Brightspace and, thereby, checked for plagiarism. If fraud or
plagiarism is detected or suspected, we will inform the Board of Examiners in the usual manner. In the event of
demonstrable fraud, the sanctions delineated in Article 5.15 of the [Education and Examination Regulations
(EER)](https://students.uu.nl/en/fss/education-and-examination-regulations-eer) will apply. 
