## Procedures {-}

------------------------------------------------------------------------------------------------------------------------

### Formatting {- #formatting}

Your assignment submissions must follow certain formatting guidelines. Failing to correctly format your submissions can
impact your grade.

#### Report {-}

You must submit your Assignment 1 & 2 reports in PDF format.

- Each report must include the following information:
    - The names of all assignment authors (i.e., all group members for Assignment 1, your name for Assignment 2).
    - The Assignment Group number (only for Assignment 1).

Your reports should include appendicies containing the lavaan syntax used to estimate each model include in the
respective report.

- Include only the lavaan syntax string and the call to the estimation function used to fit the model (e.g., `cfa()`,
`sem()`, `lavaan()`).

For example, the following snippets would be appropriate.

```{r, eval = FALSE}
cfaMod <- '
f1 =~ a1 + a2 + a3
f2 =~ b1 + b2 + b3
'

cfaOut <- cfa(cfaMod, data = dat1, std.lv = TRUE)
```

```{r, eval = FALSE}
semMod <- '
f1 =~ a1 + a2 + a3
f2 =~ b1 + b2 + b3

f1 ~ f2 + x
'

semOut <- sem(semMod, data = dat1, std.lv = TRUE)
```

The snippet below, however, contains additional unecessary commands related to data ingest/processing and summarizing
the model.

```{r, eval = FALSE}
dat0 <- readRDS("../data/dataset.rds")
dat1 <- rename(dat0, x = foo, y = bar, z = baz)

pathMod <- 'foo ~ bar + baz'
pathOut <- sem(pathOut, data = dat1)

summary(pathOut)
```

#### Syntax File {- #analysis_script}

You must also submit an executable script containing the code used for your analyses.

- If you used [**KnitR**][knitr] to embed your analysis code in your report, you may submit the source code that
generates your report (i.e., an RMD [R Markdown], QMD [Quarto], or RNW [LaTeX] file).
- Otherwise, submit a standard R script.

Clean the script before submitting.

- Remove extraneous/redundant code and comments.
   - Only include the code for analyses that appear in your report.
   - Include code for data processing/clean/wrangling.
- Order the commands to match your final analysis.
   - E.g., Data ingest $\rightarrow$ Data processing $\rightarrow$ Descriptive analysis $\rightarrow$ Modeling
   $\rightarrow$ Processing results $\rightarrow$ Post-hoc analysis
   - In theory, I should be able to source your script (i.e., run the entire script in "batch mode") to reproduce your
   results.

------------------------------------------------------------------------------------------------------------------------

### Length {-}

You may use as many words as necessary to adequately explain yourself, but concision and parsimony are encouraged. Note
that the assignments are **not** intended to be full-blown research papers! The focus should be on the definition of
your model, how this model relates to theory (introduction), and what you have learned from your estimated model
(discussion).

For both assignments, you should be able to get the job done in fewer than 10 pages of text (excluding title page,
figures, appendices, and references).

------------------------------------------------------------------------------------------------------------------------

### Submission {-}

You will submit your reports through Brightspace.

- Each assignment has a corresponding item in the "Assignments" section of the Brightspace page through which you will
submit your reports.
- For Assignment 1, you may only submit one report per group.
   - Designate one group member to submit the report.
   - The grade for this submission will apply to all group members.
- If something goes wrong with the submission, or you notice a mistake (before the deadline) that you want to correct,
you may upload a new version of your report. 
   - We will grade the final submitted version.
- The submissions will be screened with *TurnItIn* to check for plagarism.
