## At-Home Exercises

```{r include = FALSE}
source(here::here("code", "supportFunctions.R"))
```

------------------------------------------------------------------------------------------------------------------------

### Multiple-Group Path Analysis {#mgPathAnalysis}

------------------------------------------------------------------------------------------------------------------------

To fix ideas, we'll start these practical exercises by re-running part of the moderation analysis from the
[Week 3 At-Home Exercises](#moderation) as a multiple group model.

------------------------------------------------------------------------------------------------------------------------

####

Load the [*Sesam2.sav*][sesam2_data] data.

- *NOTE:* Unless otherwise specified, all analyses in Section \@ref(mgPathAnalysis) use these data.

<details>
  <summary>Click to show code</summary>

```{r}
library(haven)

# Read the data into an object called 'sesam2':
dataDir <- "data"
sesam2 <- read_sav(here::here(dataDir, "Sesam2.sav"))
```

</details>

------------------------------------------------------------------------------------------------------------------------

`VIEWCAT` is a nominal grouping variable, but it is represented as a numeric variable in the `sesam2` data. The levels
represent the following frequencies of Sesame Street viewership of the children in the data:

- `VIEWCAT = 1`: Rarely/Never
- `VIEWCAT = 2`: 2--3 times a week
- `VIEWCAT = 3`: 4--5 times a week
- `VIEWCAT = 4`: \> 5 times a week

We will use `VIEWCAT` as the grouping variable in our path model. To do so, we don't really need to convert `VIEWCAT`
into a factor, but, if we do, **lavaan** will give our groups meaningful labels in the output. That added clarity can be
pretty helpful.

------------------------------------------------------------------------------------------------------------------------

####

Convert *VIEWCAT* into a factor.

- Make sure that `VIEWCAT = 1` is the reference group.
- Assign the factor labels denoted above.

<details>
  <summary>Click to show code</summary>

```{r}
library(dplyr)

## Store the old version for checking:
tmp <- sesam2$VIEWCAT

## Convert 'VIEWCAT' to a factor:
sesam2 <- mutate(sesam2, 
                 VIEWCAT = factor(VIEWCAT,
                                  labels = c(
                                    "Rarely/never",
                                    "2-3 times per week",
                                    "4-5 times per week",
                                    "> 5 times per week"
                                    )
                                  )
                )

## Check the conversion:
table(new = sesam2$VIEWCAT, old = tmp, useNA = "always")
```

</details>

------------------------------------------------------------------------------------------------------------------------

####

Create a conditional slopes plot to visualize the effect of `AGE` on `POSTNUMB` within each of the `VIEWCAT` groups.

- Based on this visualization, do you think it is reasonable to expect that `VIEWCAT` moderates the effect of `AGE` on
`POSTNUMB`?

<details>
  <summary>Click to show code</summary>

```{r}
library(ggplot2)

ggplot(sesam2, aes(AGE, POSTNUMB, color = VIEWCAT)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE)
```

  <details>
    <summary>Click for explanation</summary>

The regression lines representing the conditional focal effects are not parallel, so there appears to be some level of
moderation. That being said, the differences are pretty small, so the moderation may not be significant (i.e., the
non-parallel regression lines may simply be reflecting sampling variability).

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

We will use path analysis to test if `VIEWCAT` moderates the effect of `AGE` on `POSTNUMB`. This analysis will entail
three steps:

1. Estimate the unrestricted multiple-group model wherein we regress `POSTNUMB` onto `AGE` and specify `VIEWCAT` as the
   grouping factor.
1. Estimate the restricted model wherein we constrain the `AGE` $\rightarrow$ `POSTNUMB` effect to be equal in all
   `VIEWCAT` groups.
1. Conduct a $\Delta \chi^2$ test to compare the fit of the two models.

------------------------------------------------------------------------------------------------------------------------

#### {#fullPaMod}

Estimate the unrestricted path model described above.

- Include the intercept term in your model.
- Judging from the focal effects estimate in each group, do you think moderation is plausible? 

<details>
  <summary>Click to show code</summary>

```{r}
library(lavaan)

## Estimate the unrestricted model a view the results:
out_full <- sem('POSTNUMB ~ 1 + AGE', data = sesam2, group = "VIEWCAT")
summary(out_full)
```

  <details>
    <summary>Click for explanation</summary>

There are some notable differences in the `AGE` $\rightarrow$ `POSTNUMB` focal effect between `VIEWCAT` groups. It looks
like `VIEWCAT` could moderate the focal effect.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

#### {#resPaMod}

Estimate the restricted model described above.

- Equate the focal effect across all `VIEWCAT` groups.

<details>
  <summary>Click to show code</summary>

```{r}
## Estimate the restricted model and view the results:
out_res <- sem('POSTNUMB ~ 1 + c("b1", "b1", "b1", "b1") * AGE',
               data = sesam2,
               group = "VIEWCAT")
summary(out_res)
```

</details>

------------------------------------------------------------------------------------------------------------------------

#### 

Test for moderation by comparing the full and restricted models from \@ref(fullPaMod) and \@ref(resPaMod).

- Does `VIEWCAT` significantly moderate the effect of `AGE` on `POSTNUMB`?

<details>
  <summary>Click to show code</summary>

```{r, paged.print = FALSE}
## Test for moderation:
anova(out_full, out_res)
```

  <details>
    <summary>Click for explanation</summary>
  
```{r, include = FALSE}
tmp <- anova(out_full, out_res)
  
chi2 <- tmp[2, "Chisq diff"] %>% round(3)
df   <- tmp[2, "Df diff"]
p    <- tmp[2, "Pr(>Chisq)"] %>% round(3)
```
  
No, *VIEWCAT* does not significantly moderate the effect of *AGE* on *POSTNUMB*
($\Delta \chi^2[`r df`] = `r chi2`$, $p = `r p`$).

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

### Multiple-Group CFA {#mgCFA}

------------------------------------------------------------------------------------------------------------------------

In the next part of these exercises, we will estimate a multiple-group CFA to evaluate the measurement structure of a
scale assessing *Prolonged Grief Disorder*. The relevant data are contained in the [*PGDdata2.txt*][pgd_data] file. This
dataset is a subset of a larger sample collected from bereaved people who had recently lost a close family member. You
can find more information about this scale in [Boelen et al. (2010)][boelen_et_al_2010].

The dataset consists of a grouping variable, `Kin2`, and 5 items taken from the *Inventory of Complicated Grief*: 

1. `Kin2`: Relation to the deceased family member {1 = Spouse, 0 = Other}
1. `b1pss1`: Yearning
1. `b2pss2`: Part of self died
1. `b3pss3`: Difficulty accepting the loss
1. `b4pss4`: Avoiding reminders of deceased
1. `b5pss5`: Bitterness about the loss 

The *PGDdata2.txt* file has the following characteristics.

- Fields are separated by tab characters.
- The first row contains variable names.
- Missing values are coded as -999.

------------------------------------------------------------------------------------------------------------------------

####

Load the *PGDdata2.txt* data.

- Use the `read.table()` function to load the data.
- Convert the missing values to `NA` via the `na.strings` argument.
- Retain the column labels via the `header` argument.
- Specify the field delimiter as the tab character (i.e., `"\t"`).

<details>
  <summary>Click to show code</summary>

```{r}
## Load the data:
pgd <- read.table(here::here(dataDir, "PGDdata2.txt"),
                  na.strings = "-999",
                  header = TRUE,
                  sep = "\t")

## Check the results:
head(pgd)
summary(pgd)
str(pgd)
```

</details>

------------------------------------------------------------------------------------------------------------------------

#### {#processPgdData}

Process the PGD data as follows.

- Convert `Kin2` into a factor variable.
- Exclude any cases with missing values on `Kin2`.
- Save the processed dataset as an RDS file.

*NOTE:* Unless otherwise specified, all analyses in Section \@ref(mgCFA) use these data.

<details>
  <summary>Click to show code</summary>

```{r}
## Process data
pgd <- mutate(pgd, Kin2 = factor(Kin2, levels = 0:1, labels = c("spouse", "other"))) |>
  filter(!is.na(Kin2))

## Save the processed data
saveRDS(pgd, here::here(dataDir, "pgd_data.rds"))
```

</details>

------------------------------------------------------------------------------------------------------------------------

#### {#oneGroupCFA}

Run a single-group CFA wherein the five scale variables described above indicate a single latent factor.

- Do not include any grouping variable.
- Use the default settings in the `cfa()` function.

<details>
  <summary>Click to show code</summary>

```{r}
## Define the model syntax:
cfaMod <- 'grief =~ b1pss1 + b2pss2 + b3pss3 + b4pss4 + b5pss5'

## Estimate the model:
out0 <- cfa(cfaMod, data = pgd)
```

</details>

------------------------------------------------------------------------------------------------------------------------

#### 

Summarize the evaluate the fitted CFA

- Does the model fit well?
- Are the items homogeneously associated with the latent factor?
- Which item is most weakly associated with the latent factor?

<details>
  <summary>Click to show code</summary>

```{r}
## Summarize the fitted model:
summary(out0, fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)
```

```{r, echo = FALSE}
fit0 <- fitMeasures(out0, c("chisq", "df", "pvalue", "rmsea", "cfi", "tli", "srmr"))

chi2 <- fit0["chisq"] %>% round(2)
df <- fit0["df"]
p <- prettyPValue(fit0["pvalue"])
rmsea <- fit0["rmsea"] %>% round(3)
cfi <- fit0["cfi"] %>% round(3)
srmr <- fit0["srmr"] %>% round(3)

r2 <- inspect(out0, "r2")
ly <- inspect(out0, "std.all")$lambda

bad <- r2 %>% which.min() %>% names()

r2 <- r2[bad] %>% round(3) %>% as.numeric()
ly <- ly[bad, ] %>% round(3) %>% as.numeric()
```

  <details>
    <summary>Click for explanation</summary>

- The model fits the data quite well
$(\chi^2[`r df`] = `r chi2`,$
$`r p`,$
$\textit{RMSEA} = `r rmsea`,$
$\textit{CFI} = `r cfi`,$
$\textit{SRMR} = `r srmr`).$
- All of the indicators appear to be more-or-less equally good indicators of the latent factor except for `r bad` which
has a standardized factor loading of $\lambda = `r ly`$ and $R^2 = `r r2`$.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

#### {#twoGroupCFA}

Rerun the CFA from \@ref(oneGroupCFA) as a multiple-group model.

- Use the `Kin2` variable as the grouping factor.
- Do not place any equality constraints across groups.

<details>
  <summary>Click to show code</summary>

```{r}
out1 <- cfa(cfaMod, data = pgd, group = "Kin2")
```

</details>

------------------------------------------------------------------------------------------------------------------------

####

Summarize the fitted multiple-group CFA from \@ref(twoGroupCFA).

- Does the two-group model fit the data well?
- Do you notice any salient differences between the two sets of within-group estimates?
  
<details>
  <summary>Click to show code</summary>

```{r}
summary(out1, fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)
```

```{r, echo = FALSE}
fit1 <- fitMeasures(out1, c("chisq", "df", "pvalue", "rmsea", "cfi", "tli", "srmr"))

chi2 <- fit1["chisq"] %>% round(2)
df <- fit1["df"]
p <- prettyPValue(fit1["pvalue"])
rmsea <- fit1["rmsea"] %>% round(3)
cfi <- fit1["cfi"] %>% round(3)
srmr <- fit1["srmr"] %>% round(3)
```

  <details>
    <summary>Click for explanation</summary>

- The two-group model also fits the data very well
$(\chi^2[`r df`] = `r chi2`,$
$`r p`,$
$\textit{RMSEA} = `r rmsea`,$
$\textit{CFI} = `r cfi`,$
$\textit{SRMR} = `r srmr`).$
- No, there are no striking differences between the two sets of estimates. Although there is certainly some variability
between groups, the two sets of estimates don't look systematically different.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

####

Based on the above results, what can you conclude about configural, weak, and strong measurement invariance across the
`Kin2` groups?

<details>
  <summary>Click for explanation</summary>

- Configural invariance holds. The unrestricted multiple-group CFA fits the data adequately (very well, actually), and
the measurement model parameters are reasonable in both groups.
- We cannot yet draw any conclusions about weak or strong invariance. We need to do the appropriate model comparison
tests first.

</details>

------------------------------------------------------------------------------------------------------------------------

End of At-Home Exercises

------------------------------------------------------------------------------------------------------------------------

[sesam2_data]: https://surfdrive.surf.nl/files/index.php/s/dfzC7Tf5HHiTX8M/download
[pgd_data]: https://surfdrive.surf.nl/files/index.php/s/xxkp4gZY682AGyY/download
[boelen_et_al_2010]: https://doi.org/10.1016/j.jad.2010.01.076
