## At-Home Exercises

```{r include = FALSE}
opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE)
dataDir <- "data"
```

### Setup

To complete the following exercises, you'll need to install the packages described in the following table.

```{r, echo = FALSE}
data.frame(
  Package = c(
    "haven",
    "tidySEM"
  ),
  Description = c(
    "Loading data from SPSS '.sav' files",
    "Plotting and tabulating the output of SEM-models"
  )
) |> 
  knitr::kable()
```

Use the `install.packages()` function to install these packages.

<details>
  <summary>Click for explanation</summary>

1. Open RStudio
1. Inside RStudio, find the window named **Console** on left side of the screen.
3. Copy the following code into the console and hit `Enter`/`Return` to run the command.

```{r, eval = FALSE}
install.packages(c("haven", "tidySEM"), dependencies = TRUE)
```

</details>

------------------------------------------------------------------------------------------------------------------------

###

Use the `haven::read_spss()` function to load the `LifeSat.sav` data. 

```{r, echo = FALSE}
library(dplyr)
library(haven)
LifeSat <- read_spss(here::here(dataDir, "LifeSat.sav"))
```

<details>
  <summary>Click for explanation</summary>

```{r, eval = FALSE}
library(dplyr)
library(haven)

LifeSat <- read_spss("LifeSat.sav") 
```

</details>

------------------------------------------------------------------------------------------------------------------------

###

Make a table of descriptive statistics for the variables: `LifSat`, `educ`, `ChildSup`, `SpouSup`, and `age`. 

- What is the average age in the sample? 
- What is the range (youngest and oldest child)? 

*Hint:* Use the `tidySEM::descriptives()` function.`

<details>
  <summary>Click for explanation</summary>

- The package `tidySEM` contains the `descriptives()` function for computing descriptive statistics. 
- The `describe()` function in the `psych` package is a good alternative.

```{r}
library(tidySEM)

descriptives(LifeSat[ , c("LifSat", "educ", "ChildSup", "SpouSup", "age")])
```
</details>

------------------------------------------------------------------------------------------------------------------------

### {#slr1}

Run a simple linear regression with `LifSat` as the dependent variable and `educ` as the independent variable.

*Hints:*

- The `lm()` function (short for linear model) does linear regression. 
- The `summary()` function provides relevant summary statistics for the model. 
- It can be helpful to store the results of your analysis in an object.

<details>
  <summary>Click for explanation</summary>

```{r}
results <- lm(LifSat ~ educ, data = LifeSat)
summary(results)
```
</details>

------------------------------------------------------------------------------------------------------------------------

### {#slr2}

Repeat the analysis from \@ref(slr1) with `age` as the independent variable. 

<details>
  <summary>Click for explanation</summary>

```{r}
results <- lm(LifSat ~ age, data = LifeSat)
summary(results)
```
</details>

------------------------------------------------------------------------------------------------------------------------

### {#slr3}

Repeat the analysis from \@ref(slr1) and \@ref(slr2) with `ChildSup` as the independent variable. 

<details>
  <summary>Click for explanation</summary>

```{r}
results <- lm(LifSat ~ ChildSup, data = LifeSat)
summary(results)
```
</details>

------------------------------------------------------------------------------------------------------------------------

### {#mlr}

Run a multiple linear regression with `LifSat` as the dependent variable and `educ`, `age`, and `ChildSup` as the
independent variables. 

*Hint:* You can use the `+` sign to add multiple variables to the right hand side (RHS) of your model formula.

<details>
  <summary>Click for explanation</summary>

```{r}
results <- lm(LifSat ~ educ + age + ChildSup, data = LifeSat)
summary(results)
```
</details>

------------------------------------------------------------------------------------------------------------------------

### 

Compare the results from \@ref(mlr) with those from \@ref(slr1), \@ref(slr2), and \@ref(slr3).

- What do you notice when you compare the estimated slopes for each of the three predictors in the multiple regression
model with the corresponding estimates from the simple regression models? 

------------------------------------------------------------------------------------------------------------------------

End of At-Home Exercises

------------------------------------------------------------------------------------------------------------------------
