## At-Home Exercises

```{r, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(error = TRUE)
source(here::here("code", "supportFunctions.R"))
```

This week, we'll take another look at the Kestilä (2006) results. During this practical, you will use SEM to conduct a 
regression analysis of the *Trust in Politics* factors that you estimated in the
[Week 5 At-Home Exercises](at-home-exercises-4.html).

------------------------------------------------------------------------------------------------------------------------

###

Load the Finnish subsample of ESS data.

- The relevant data are contained in the [*ess_finland.rds*][fin_data] file.
- This dataset comprises the Finnish subsample from the Week 5 exercise data.
   - The missing values have been imputed with a single round of predictive mean matching.

*Note:* Unless otherwise noted, all the following analyses use these data.

<details>
  <summary>Click to show code</summary>
  
```{r}
dataDir <- "data"
ess_fin <- readRDS(here::here(dataDir, "ess_finland.rds"))
```

</details>

------------------------------------------------------------------------------------------------------------------------

We need to do a little data processing before we can fit the regression model. At the moment, `lavaan` will not
automatically convert a factor variable into dummy codes. So, we need to create explicit dummy codes for the two factors
we'll use as predictors in our regression analysis: *sex*  and *political orientation*.  

------------------------------------------------------------------------------------------------------------------------

###

Convert the *sex* and *political interest* factors into dummy codes.

<details>
  <summary>Click to show code</summary>
  
```{r}
library(dplyr)

## Create a dummy codes by broadcasting a logical test on the factor levels:
ess_fin <- mutate(ess_fin,
                  female = ifelse(sex == "Female", 1, 0),
                  hi_pol_interest = ifelse(polintr_bin == "High Interest", 1, 0)
                  )

## Check the results:
with(ess_fin, table(dummy = female, factor = sex))
with(ess_fin, table(dummy = hi_pol_interest, factor = polintr_bin))
```
  <details>
    <summary>Click for explanation</summary>

In R, we have several ways of converting a factor into an appropriate set of dummy codes.

- We could use the `dplyr::recode()` function.
- We can use the `model.matrix()` function to define a design matrix based on the inherent contrast attribute of the
factor.
    - Missing data will cause problems here.
- We can us `as.numeric()` to revert the factor to its underlying numeric representation {Male = 1, Female = 2} and use
arithmetic to convert {1, 2} $\rightarrow$ {0, 1}.

When our factor only has two levels, though, the `ifelse()` function is the simplest way.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

We are now ready to estimate our latent regression model. Specifically, we want to estimate a single SEM to regress the
three *Trust in Politics* factors onto the following predictors:

- *Sex*
- *Age*
- *Years of Education*
- *High vs. Low Political Interest*
- *Placement on the Left-Right Political Scale*

The following path diagram shows the intended theoretical model.

```{r echo = FALSE}
library(semPlot)

mod_sem <- '
## Define the latent DVs:
institutions =~ trstlgl + trstplc + trstun + trstep + trstprl
satisfaction =~ stfhlth + stfedu  + stfeco + stfgov + stfdem
politicians  =~ pltinvt + pltcare + trstplt

## Specify the structural relations:
institutions +
satisfaction +
politicians ~ female + age + eduyrs + hi_pol_interest + lrscale
'

sem(mod_sem, data = ess_fin, fixed.x = FALSE) %>%
  semPaths(layout = "tree2",
           rotation = 2,
           residuals = FALSE,
           sizeMan = 10,
           sizeMan2 = 3,
           sizeLat = 15,
           sizeLat2 = 8,
           curve = 2,
           esize = 1,
           asize = 2,
           cardinal = "manifest",
           nCharNodes = 0)
```

Although the variances are not included in this path diagram, all variables in the model (including the observed
predictor variables) are random.

------------------------------------------------------------------------------------------------------------------------

###

Define the `lavaan` model syntax for the SEM shown above.

- Use the definition of the `institutions`, `satsifaction`, and `politicians`
  factors from \@ref(cfaSyntax) to define the DVs.
- Covary the three latent factors.
- Covary the five predictors.

<details>
<summary>Click to show code</summary>

```{r}
mod_sem <- '
## Define the latent DVs:
institutions =~ trstlgl + trstplc + trstun + trstep + trstprl
satisfaction =~ stfhlth + stfedu  + stfeco + stfgov + stfdem
politicians  =~ pltinvt + pltcare + trstplt

## Specify the structural relations:
institutions ~ female + age + eduyrs + hi_pol_interest + lrscale
satisfaction ~ female + age + eduyrs + hi_pol_interest + lrscale
politicians  ~ female + age + eduyrs + hi_pol_interest + lrscale
'
```

<details>
  <summary>Click for explanation</summary>

- We simply need to add a three lines defining the latent regression paths to our old CFA syntax.
- We don't need to specify the covariances in the syntax. We can use options in the `sem()` function to request those
estimates.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

### {#semEstimation}

Estimate the SEM, and summarize the results.

- Fit the model to the processed Finnish subsample from above.
- Estimate the model using `lavaan::sem()`.
- Use the fixed-factor method to identify the model.
- Request the standardized parameter estimates with the summary.
- Request the $R^2$ estimates with the summary.

<details>
  <summary>Click to show code</summary>

```{r}
library(lavaan)

## Fit the SEM:
fit_sem <- sem(mod_sem, data = ess_fin, std.lv = TRUE, fixed.x = FALSE)

## Summarize the results:
summary(fit_sem, fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)
```

<details>
  <summary>Click for explanation</summary>

The `fixed.x = FALSE` argument tells **lavaan** to model the predictors as random variables. By default, **lavaan** will
covary any random predictor variables. So, we don't need to make any other changes to the usual procedure.

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

In the final part of these exercises, you will rerun the above SEM as a path model wherein the observed mean scores
representing *Trust in Institutions*, *Satisfaction with Political Systems*, and *Trust in Politicians* act as the DVs.
The `scoreItems()` function from the **psych** package makes it very easy to create scale scores.

------------------------------------------------------------------------------------------------------------------------

###

Read the documentation for the `psych::scoreItems()` function.


<details>
  <summary>Click to show code</summary>

```{r, eval = FALSE}
?psych::scoreItems
```

</details>

------------------------------------------------------------------------------------------------------------------------

###

Use the `psych::scoreItems()` function to create mean scores to represent the three latent constructs in the above SEM.

<details>
  <summary>Click to show code</summary>

```{r}
library(psych)

## Create a list to define the item-to-score mapping
scaleKeys <- list(
  ins_score = c("trstlgl", "trstplc", "trstun", "trstep", "trstprl"),
  sat_score = c("stfhlth", "stfedu", "stfeco", "stfgov", "stfdem"),
  pol_score  = c("pltinvt", "pltcare", "trstplt")
  )

## Create the scale scores
scaleScores <- scoreItems(keys = scaleKeys, items = ess_fin, impute = "none")

## If we print the 'scaleScores' object, we get a bunch of interesting psychometric information
scaleScores

## Add the scores to the ESS dataset
ess_fin <- data.frame(ess_fin, scaleScores$scores)
```

</details>

------------------------------------------------------------------------------------------------------------------------

###

Use the mean scores you created above to estimate a path analytic version of the SEM from \@ref(semEstimation).

<details>
  <summary>Click to show code</summary>

```{r}
## Specify the model syntax for the path model:
mod_pa <- '
## Specify the structural relations:
ins_score ~ female + age + eduyrs + hi_pol_interest + lrscale
sat_score ~ female + age + eduyrs + hi_pol_interest + lrscale
pol_score ~ female + age + eduyrs + hi_pol_interest + lrscale
'

## Fit the path analytic model:
fit_pa <- sem(mod_pa, data = ess_fin, fixed.x = FALSE)

## Summarize the results:
summary(fit_pa, fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)
```

</details>

------------------------------------------------------------------------------------------------------------------------

###

Compare the results from the path analysis to the SEM-based results.

- Does it matter whether we use a latent variable or a factor score to define the DV?

*Hint:* When comparing parameter estimates, use the fully standardized estimates (i.e., the values in the column labeled
`Std.all`).

<details>
  <summary>Click to show code</summary>

*Note:* The "supportFunction.R" script that we source below isn't a necessary part of the solution. This script defines
a bunch of convenience functions. One of these functions, `partSummary()`, allows us to print selected pieces of the
model summary.

```{r, eval = FALSE}
## Source the script that defines the partSummary() convenience function:
source(here::here("code", "supportFunctions.R"))
```

```{r}
## View the regression estimates from the SEM:
partSummary(fit_sem, 8, standardized = TRUE)

## View the regression estimates from the path analysis:
partSummary(fit_pa, 7, standardized = TRUE)

## View the R-squared estimates from the SEM:
partSummary(fit_sem, 11, rsquare = TRUE)

## View the R-squared estimates from the SEM:
partSummary(fit_pa, 10, rsquare = TRUE)
```

```{r, eval = FALSE, echo = FALSE}
r2_sem <- summary(fit_sem, rsquare = TRUE)$pe %>% 
  filter(op == "r2", rhs == "institutions") %>% 
  select(est) %>% 
  as.numeric()
r2_pa <- inspect(fit_pa, "r2")
```

  <details>
    <summary>Click for explanation</summary>

```{r, echo = FALSE, include = FALSE}
estSem <- parameterEstimates(fit_sem, standardized = TRUE)
estPa  <- parameterEstimates(fit_pa, standardized = TRUE)

x <- estSem |> filter(op == "~") |> select(lhs, rhs, std.all) |> rename(X = rhs, SEM = std.all)
y <- estPa |> filter(op == "~") |> select(std.all) |> rename(PA = std.all)

beta <- data.frame(x, y)

betaIns <- filter(beta, lhs == "institutions") |> select(-lhs)
betaSat <- filter(beta, lhs == "satisfaction") |> select(-lhs)
betaPol <- filter(beta, lhs == "politicians") |> select(-lhs)

xx <- estSem |>
  filter(op == "~") |>
  select(lhs, rhs, est, se, z) |>
  rename(X = rhs, SEM_Est = est, SEM_SE = se, SEM_Z = z)
yy <- estPa |>
  filter(op == "~") |>
  select(est, se, z) |>
  rename(PA_Est = est, PA_SE = se, PA_Z = z)

tests <- data.frame(xx, yy)

testIns <- filter(tests, lhs == "institutions") |> select(-lhs)
testSat <- filter(tests, lhs == "satisfaction") |> select(-lhs)
testPol <- filter(tests, lhs == "politicians") |> select(-lhs)

r2 <- data.frame(
  SEM = lavInspect(fit_sem, "r2") |> tail(3), 
  PA = lavInspect(fit_pa, "r2") |> as.numeric()
)
```

The two approaches reach the same substantive conclusions, but we do see differences. For example, the standardized
regression weights mostly suggests larger effect sizes in the SEM model.

#### DV = Institutions {-}

```{r, echo = FALSE}
kable(betaIns, digits = 3)
```

#### DV = Satisfaction {-}

```{r, echo = FALSE}
kable(betaSat, digits = 3)
```
  
#### DV = Politicians {-}

```{r, echo = FALSE}
kable(betaPol, digits = 3)
```

The $R^2$ values from the SEM are also larger than those from the path analysis for two out of three factors and
approximately equal for the *Politicians* factor.

```{r, echo = FALSE}
kable(r2, digits = 3)
```

  </details>
</details>

------------------------------------------------------------------------------------------------------------------------

End of At-Home Exercises

------------------------------------------------------------------------------------------------------------------------

[fin_data]: https://surfdrive.surf.nl/files/index.php/s/rN8FuscpJTLsB4h/download
